---
title: 'Project: TVL of DeFi Protocols'
author: "Francesco Decataldo"
date: "15/06/2021"
output:
  ioslides_presentation: default
  slidy_presentation: default
  beamer_presentation: default
---

<style>
slides > slide {
  overflow-x: auto !important;
  overflow-y: auto !important;
}

slides > slide:not(.nobackground):after {
  content: '';
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(cowplot)
library(shiny)
library(dygraphs)
```

```{r Load data, include=FALSE}
protocols = readRDS("protocols.rds")
plots = readRDS("plots.rds")
aprPlots = readRDS("aprPlots.rds")
```



## Overview {.build}

* In this presentation we perform a study of some **DeFi platforms**; in particular we focus on a metric, the **TVL (Total Value Locked)**, an index of the capitalization and liquidity of a certain platform. According to the [DYOR Crypto wiki](https://dyor-crypto.fandom.com/wiki/Total_Locked_Value_(TLV)) it is defined as follows:

> TVL represents the amount of assets that are currently being staked in a specific protocol. This value is by no means meant to represent the amount of outstanding loans, but rather the total amount of underlying supply being secured by a specific application and/or by DeFi as a whole.

* After a general analysis, we focus on a study of the **APR (Annual Percentage Rate)**  of the liquidity mining program of **Pancake Swap**.

## TVL {.build}

Which are the most capitalized platform, based on the **TVL**?

```{r echo=FALSE}
plots[[1]]
```


```{r echo=FALSE}
plot_grid(plots[[2]],
plots[[3]],
plots[[4]], ncol=3)
```

What we observe is that:

* **Ethereum** is the blockchain where most platforms have locked tokens.
* On **Binance Smart Chain**, **PancakeSwap** is the only platform with a comparable TVL.
* **Polygon** is almost used by the same platforms as **Ethereum**  (apart from **Quickswap**).


## TVL over time {.build}

Analysis of the 5 platforms with higher capitalization, starting from beginning of 2021:

1. AAVE 
2. Curve
3. Pancake Swap
4. Compound
5. Uniswap

```{r Plot1, echo=FALSE}
do.call("plot_grid", c(protocols$plot1, ncol=3))
```

* Similar pattern in the **TVL**: is there a **common factor** affecting **TVL**?

## TVL vs ETH {.build}

we compare the TVL with the price of **Ether(ETH)**, the token on which most of our platforms are based:

> Click [this link](https://fradeca01.shinyapps.io/InteractiveTVL/) for an interactive plot.

```{r}
do.call("plot_grid", c(protocols$plot2, ncol=3))
```

* Correlation between **TVL** and **ETH**

Scatter plot with **TVL** against **ETH Price**:

```{r}
do.call("plot_grid", c(protocols$plot3, ncol=3))
```

## Modelling {.build}

Now, we build a model for each platform, assuming a linear relation:

```{r}
do.call("plot_grid", c(protocols$plot4, ncol=3))
```

```{r}

r2 = protocols$r2

names(r2) = protocols$name

r2
```

## Pancake Swap {.build}

**PancakeSwap** is developed on the **Binance Smart Chain** and it does not have **ETH** as main token; why the correlation with **ETH** is good anyway? 

<!-- Correlation between **BNB** and **ETH** prices: -->

<!-- ```{r echo=FALSE} -->
<!-- plots[[5]] -->
<!-- ``` -->

<!-- * **ETH** and the **BNB** are correlated  -->

Model of **PancakeSwap** **TVL** with **BNB Price**:

```{r echo=FALSE}
plots[[6]]
```

At this point we have learned that:

$$
\small TVL \propto \text{CHAIN TOKEN}
$$

## PancakeSwap APR {.build .smaller}

What is **APR**?

>The **APR(Annual Percentage Rate)** represents the yield obtained from staking funds on a certain pool (without considering the composability).

It's calculated with the following formula (as we can see [here](https://github.com/pancakeswap/pancake-frontend/blob/develop/src/utils/apr.ts)):

$$
\small APR = \text{CAKES PER YEAR}\times \frac{\text{POOL WEIGHT}}{\text{TOTAL LIQUIDITY}} \times \text{CAKE PRICE}
$$
Considering that **CAKES PER YEAR** and **POOL WEIGHT** are constants we could rewrite the formula as:

$$
\small APR = k\,\,\frac{\text{CAKE PRICE}}{\text{TOTAL LIQUIDITY}}
$$
In the above slides we see that the **TVL** of **PancakeSwap** is approximately proportional to the price of **BNB**, this yield:


$$
\small APR \,\propto \,\frac{CAKE}{BNB}
$$

## CAKE vs BNB {.build}

Is there a correlation between **CAKE** and **BNB** prices?

```{r}
aprPlots[[1]]
```

The **BNB** price and the **CAKE** price has been strongly correlated during last year,  i.e. $\small CAKE =k\, BNB$. Plugging this result in last formula we obtain that the **APR** should has been approximately constant.

## APR is actually costant? {.build}

Let's verify this, plotting the **APR** of some **PancakeSwap** pools:

```{r message=FALSE, warning=FALSE}
do.call("plot_grid", aprPlots[-1])
```



